---
title: "R Markdown and Plotly"
author: "Shivanand R Koppalkar"
date: "April 28, 2017"
output: html_document
---

## Background

* As of 2014, about 4 million people retire every year.
* Some states are more "Retirement Friendly" than others.
* In Mar 2015, "BankRate" published a "First to Worst" survey, ranking states according to:
    + Cost of Living
    + Crime Rate
    + Health Care Quality
    + Weather
    + Community Well-being
* According to the BankRate survey, 60% of retirees will move when they retire.

## Obtaining the Data

* We "scraped" the "First to Worst" data from the "BankRate" website at: 
  "http://www.bankrate.com/finance/retirement/best-places-retire-how-state-ranks.aspx""

* After minor data cleanup, and adding the standard Postal Code State Abbreviations, 
  we were ready to visualize the data.


# Load the Plotly, xlsx and Webshot Libraries
* We loaded the required libraries as needed: 
```{r setup, echo=TRUE, message=FALSE}
if (!require("plotly")) install.packages("plotly")
library(plotly)
if (!require("webshot")) install.packages("webshot")
library(webshot)
if (!require("xlsx")) install.packages("xlsx")
library("xlsx")
```

# Load the Bank Rate Survey File
```{r getData, echo=FALSE}
brFileName <- "C:/2015.03.BankRateSurvey.xlsx"
if (!file.exists(brFileName)) stop ("Data File is Missing")
brData <- read.xlsx(brFileName, sheetIndex = 1, header = TRUE)
brSiz <- dim(brData)
brData$Overall.rank <- as.vector(brData$Overall.rank)
```

* We loaded the `r brSiz[1]` row x `r brSiz[2]` column data frame from `r brFileName`
* The data column headers are: `r names(brData)`
    * Each column contains a "First to Worst" ranking for that category; low numbers are good;
    * The "Overall" ranking is according to BankRate's weighting criteria; a future version will support custom criteria.

```{r doData, echo=FALSE}
# Create hover text
brData$hover <- with(brData, paste(state.name,
                                   '<br>', "Overall Rank: ", Overall.Rank,
                                   '<br>', "Cost of Living: ", Cost.of.Living,
                                   '<br>', "Crime Rate: ", Crime.Rate,
                                   '<br>', "Well Being: ", Community.Well.being,
                                   '<br>', "Health Care: ", Health.Care,
                                   '<br>', "Tax Rate: ", Tax.Rate,
                                   '<br>', "Weather: ", Weather))

# Make state borders black
borders <- list(color = toRGB("black"), width=2)

# Set up some mapping options
map_options <- list(scope = 'usa', 
                    projection = list(type = 'albers usa'),
                    showlakes = TRUE,
                    lakecolor = toRGB('black')
                    )
```
* We set parameters for the Choropleth map, and assembled information to be displayed   
  whenever the mouse hovers over a state; may need to zoom the map to see it all.           
## First to Worst Retirement Friendly States

```{r plotData, echo=FALSE}
plotly::plot_ly(brData, 
                z = brData$Overall.Rank, 
                text = brData$hover, 
                locations = brData$State, 
                type = 'choropleth', 
                locationmode = 'USA-states', 
                color = brData$Overall.Rank, 
                colors = 'Reds', 
                marker = list(line = borders)) %>%
  colorbar(title = "Retiree<br>Hostility") %>%
  layout(title = 'Best States to Retire<br>(Hover for Details)', 
         geo = map_options)                    
```         